Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-03-02 src/article.tex
Proof read new sections

These methods provide a wide variety of possible approaches for scaffolding
a genome. The scaffold may still however require manual editing and addition of
sequence if computational methods are not completely successful. This involves
moving and joining larges blocks of nucleotide text together by hand which may
be error-prone and precludes any reproducibility. \pb

\subsection*{Overview} %{{{2

The software outlined here, ``Scaffolder'', is aimed at creating a syntax and
software framework for producing and editing a genome scaffold. Scaffolder uses
a specific file format to define how contigs are joined, additional sequences
inserted, and for the specification of unknown regions. This syntax allows
a scaffold to be updated by simply editing the scaffold file and allows the
scaffolding process to be both reproducible and deterministic. \pb

@@

\subsection{Scaffolder Sofware Interface} %{{{2

Scaffolder provides a standardised set of Ruby classes and methods (API) for
interacting with the scaffold. This allows Scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as Rake. In
addition Scaffolder provides a command line interface (CLI) providing functions
to validate the scaffold file and build a draft super sequence. This allows the
scaffold sequence to be generated and validated without any Ruby experience.
The Scaffolder CLI behaves as a standard Unix tool and returns appropriate exit
codes and manual pages. The use of both these Scaffolder interfaces is outlined
in detail on the scaffolder website (\scaffolder). \pb

@@

Scaffolding is a intermediate stage in a genome project. Additional sequences
are produced using a combination of computational and laboratory methods. These
new sequences are then added to existing sequencing data to produce
a continuous sequence of the target organism genome. The scaffolding process
however is a laborious manual process involving joining and adding large
regions of nucleotide sequence together. Any manual process is inherently hard
to reproduce and introduces the possibility of human error. The aims of
Scaffolder are therefore twofold. The first is providing software that is easy
to install and simplifies the task of genome finishing. The second aim is to
facilitate reproducibility when joining large nucleotide sequences together to
create a genome sequence. \pb

@@

Scaffolder is software aimed at bioinformaticians and biologists familar with
the command line who wish to build scaffold a genome sequence in a concise
readable way. The Scaffolder file format allows other biologists to see how the
genome sequence was assembled and allows a broad overview of how the sequence
was assembled not possible not possible from a megabase-sized string of
nucleotide characters. Scaffolder also furthers the cause for reporducibility
in genome sequence by allowing the scaffold super-sequence to be reliably
reproduced from the scaffold file. \pb

The design of the Scaffolder software follows the Unix philosophy of ``Do one
thing and do it well.''. The focus of Scaffolder is a simple to use tool for
creating a genome scaffold. The YAML data format for creating the scaffold file
is standardised and easily manipulated programmatically. This thereby means the
scaffolding process follows a second Unix tenet: ``If your data structures are
good enough, the algorithm to manipulate them should be trivial.''.

>>>>

These methods provide a wide array of approaches for merging contigs into
larger a continuous scaffolds. The scaffolding process may still however
require inserting additional sequence or joining contigs using PCR-derived
information from the laboratory. This manual stage in the finishing process
will moving and editing large blocks of nucleotide text by hand which is subject
to human error-prone and precludes any reproducibility for the steps taken. \pb

The software outlined here, ``Scaffolder'', aims to solve address these
problems by creating a file syntax and software framework for editing a genome
scaffold. Scaffolder uses a specific file format to define how contigs are
joined, additional sequences inserted, and for the specification of unknown
regions. This syntax allows a scaffold to be updated by simply editing the
scaffold file facilitating a reproducible scaffolding process and providing
a concise overview of how the scaffold was constructed. \pb

@@

\subsection{Scaffolder Software Interface} %{{{2

Scaffolder provides a standardised set of Ruby classes and methods (API) for
interacting with the scaffold. This allows Scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as Rake. In
addition Scaffolder provides a command line interface (CLI) providing functions
to validate the scaffold file and build a draft super sequence. The Scaffolder
CLI behaves as a standard Unix tool and returns appropriate exit codes and
manual pages. The use of both these Scaffolder interfaces is outlined in detail
on the scaffolder website (\scaffolder). \pb

@@

Scaffolding requires joining assembled contigs and insert additional sequence
data using a combination of computational and laboratory methods. The process
of manually editing a scaffold is inherently hard to reproduce and introduces
the possibility of human error. The aims of this Scaffolder software are
therefore twofold. The first is providing software that is easy to install and
simplifies the task of genome finishing. The second aim is to facilitate
reproducibility when joining large nucleotide sequences together to create
a genome sequence. \pb

@@

Scaffolder is software aimed at bioinformaticians and biologists familiar with
the command line building genome sequence scaffold. The file format maintains
the scaffold in a concise and readable way allowing third parties to see how
the genome sequence was assembled. This file format allows a broad overview of
of which sequences were included and how they are ordered possible not possible
from a megabase-sized strings of nucleotide characters. Scaffolder also
furthers the cause for reproducibility in genome projects by allowing the
scaffold super-sequence to be reliably reproduced from the scaffold file. The
YAML syntax for writing the scaffold file is also standardised and to
manipulate programmatically. This thereby means the scaffolding process follows
the Unix tenet of: ``If your data structures are good enough, the algorithm to
manipulate them should be trivial.''.
